CYO Project Report - Brazilian Ecommerce

### 2021-07-06 draftReport.Rmd
updated for scaling and one hot encoding

```{r Install or load packages,include=FALSE}
# TODO use conditional loading or installation of library
# Remove plotly related code for pdfs
if (!require('dplyr')) install.packages('dplyr'); library('dplyr')
if (!require('lubridate')) install.packages('lubridate'); library('lubridate')
if (!require('tidyverse')) install.packages('tidyverse'); library('tidyverse')
if (!require('geosphere')) install.packages('geosphere'); library('geosphere')
if (!require('caret')) install.packages('caret'); library('caret')
if (!require('plotly')) install.packages('plotly'); library('plotly')
if (!require('gam')) install.packages('gam'); library('gam')
```

@@

# let's combine with products data as well
orders_dataset %>% inner_join(order_items_dataset,by='order_id') %>% inner_join(sellers_dataset,by='seller_id') %>% inner_join(customers_dataset,by='customer_id') %>% inner_join(products_dataset,by='product_id') %>% select(-customer_id) %>% head()

# seller to customer distance should be right predictor to consider, as the text says "Those merchants are able to sell their products through the Olist Store and ship them directly to the customers using Olist logistics partners."
# source: https://www.kaggle.com/olistbr/brazilian-ecommerce
# we will use distGeo from geosphere package to compute distance in km between seller and customer. considered to be highly accurate. if it takes more time, we can consider using distVincentyEllipsoid or distHaversine as we are fine with accuracy of a km.
orders_dataset %>% inner_join(order_items_dataset,by='order_id') %>% inner_join(sellers_dataset,by='seller_id') %>% inner_join(customers_dataset,by='customer_id') %>%  inner_join(products_dataset,by='product_id') %>% select(-customer_id) %>% head() %>% mutate(distance_between_two_points=round(distGeo(matrix(c(geolocation_lng_seller,geolocation_lat_seller),ncol=2),matrix(c(geolocation_lng_customer,geolocation_lat_customer),ncol=2))/1000,2))

# so raw data
raw_data <- orders_dataset %>% inner_join(order_items_dataset,by='order_id') %>% inner_join(sellers_dataset,by='seller_id') %>% inner_join(customers_dataset,by='customer_id') %>% inner_join(products_dataset,by='product_id') %>%  mutate(distance_between_two_points=round(distGeo(matrix(c(geolocation_lng_seller,geolocation_lat_seller),ncol=2),matrix(c(geolocation_lng_customer,geolocation_lat_customer),ncol=2))/1000,2)) %>% select(-geolocation_lat_customer,-geolocation_lng_customer,-geolocation_lat_seller,-geolocation_lng_seller,-customer_id,-order_status,-seller_city,-seller_state,-customer_city,-customer_state)

p1 <- raw_data %>% ggplot(aes(distance_between_two_points)) + geom_histogram(binwidth=50)
ggplotly(p1)

# TODO we maybe able to remove this as it is not used, because this is the info which is generally not available
# identify if there was delay in delivery to carrier
raw_data <- raw_data %>% mutate(delayed_to_carrier=if_else((as.numeric(handed_over_to_carrier_date)-as.numeric(shipping_limit_date))>0,TRUE,FALSE))

wrangled_data <- raw_data %>% select(-shipping_limit_date,-handed_over_to_carrier_date,-payment_approved_at,-order_id,-order_estimated_delivery_date,-order_item_id,-delayed_to_carrier) %>% mutate(order_purchase_timestamp=as.numeric(order_purchase_timestamp),order_delivered_customer_date=as.numeric(order_delivered_customer_date),product_id=as.factor(product_id),seller_id=as.factor(seller_id),seller_zip_code_prefix=as.factor(seller_zip_code_prefix),customer_unique_id=as.factor(customer_unique_id),customer_zip_code_prefix=as.factor(customer_zip_code_prefix),product_category_name=as.factor(product_category_name))

# we will predict number of hours to deliver based on various predictors. This seems more appropriate. Though as a result of this, we won't be able to use night, day time data of purchase which may influence delivery time (say to carrier it is informed only next day)
wrangled_data <- wrangled_data %>% mutate(number_of_hours=(order_delivered_customer_date-order_purchase_timestamp)/3600) %>% select(-order_delivered_customer_date,-order_purchase_timestamp) 
```

@@

```{r Cleanup and save required dataset}
save(wrangled_data,file='rda/wrangled_data.rda')
rm(geolocation_dataset, order_items_dataset, orders_dataset, p1, product_category_name_translation, products_dataset, sellers_dataset, sellers_to_ignore, customers_dataset, customers_to_ignore)
```

@@

```{r Explore product_category_name relation with number_of_hours, echo=FALSE}
wrangled_data %>% group_by(product_category_name) %>% summarise(mean=mean(number_of_hours),se=sd(number_of_hours)/sqrt(n())) %>% mutate(product_category_name=reorder(product_category_name,mean)) %>% ggplot(aes(x=product_category_name,y=mean,ymax=mean+2*se,ymin=mean-2*se)) + geom_point()+geom_errorbar() + theme(axis.text.x = element_text(angle=90))
```

@@

```{r Explore price relation with number_of_hours, echo = FALSE}
# first let see how are the items associated with orders priced. Most of the order items a
wrangled_data %>% mutate(price_group=round(price/100)+1) %>% ggplot(aes(price_group)) + geom_histogram(binwidth=1) + xlab('Each price_group is of 100 units each')

# let's plot errorbar for number of hours vs price_group.
# we can see while for price groups number of hours for delivery vary but the relationship is not clear enough. Higher priced items can also be delivered early
wrangled_data %>% mutate(price_group=round(price/50)+1) %>% mutate(price_group=as.factor(price_group)) %>% group_by(price_group) %>% summarise(mean=mean(number_of_hours),se=sd(number_of_hours)/sqrt(n())) %>% ggplot(aes(x=price_group,y=mean,ymax=mean+2*se,ymin=mean-2*se)) + geom_point()+geom_errorbar() + theme(axis.text.x = element_text(angle=90))

cor(wrangled_data$price,wrangled_data$number_of_hours)
```

@@

```{r Explore distance relationship between two points with number_of_hours, echo = FALSE}
# let's plot errorbar for number of hours vs distance group
# we can see a pattern that as distance increases delivery time increases. 
wrangled_data %>% mutate(distance_group=round(distance_between_two_points/50)+1) %>% mutate(distance_group=as.factor(distance_group)) %>% group_by(distance_group) %>% summarise(mean=mean(number_of_hours),se=sd(number_of_hours)/sqrt(n())) %>% ggplot(aes(x=distance_group,y=mean,ymax=mean+2*se,ymin=mean-2*se)) + geom_point()+geom_errorbar() + theme(axis.text.x = element_text(angle=90))

cor(wrangled_data$distance_between_two_points,wrangled_data$number_of_hours)
```

@@

```{r Explore freight value group with number_of_hours}
# let's plot errorbar for number of hours vs freight value group
wrangled_data %>% mutate(freight_value_group=round(freight_value/5)+1) %>% mutate(freight_value_group=as.factor(freight_value_group)) %>% group_by(freight_value_group) %>% summarise(mean=mean(number_of_hours),se=sd(number_of_hours)/sqrt(n())) %>% ggplot(aes(x=freight_value_group,y=mean,ymax=mean+2*se,ymin=mean-2*se)) + geom_point()+geom_errorbar() + theme(axis.text.x = element_text(angle=90))

cor(wrangled_data$freight_value,wrangled_data$number_of_hours)
```

@@

However, it dips near the end
```{r smooth fit for freight value, echo=FALSE}
wrangled_data %>% ggplot(aes(x=freight_value,y=number_of_hours)) + geom_smooth() 
```

@@

```{r Explore seller_zip_code_prefix with number_of_hours, echo = FALSE}
# TODO anyway to add regularization or remove the low impact seller_zip_code_prefix from the model
# let's plot errorbar for number of hours vs seller_zip_code_prefix
sellers_zip_of_interest <- wrangled_data %>% dplyr::count(seller_zip_code_prefix) %>% filter(n>10) %>% pull(seller_zip_code_prefix)

wrangled_data %>% filter(seller_zip_code_prefix %in% sellers_zip_of_interest) %>% group_by(seller_zip_code_prefix) %>% summarise(mean=mean(number_of_hours),se=sd(number_of_hours)/sqrt(n())) %>% mutate(seller_zip_code_prefix=reorder(seller_zip_code_prefix,mean)) %>%  ggplot(aes(x=seller_zip_code_prefix,y=mean,ymax=mean+2*se,ymin=mean-2*se)) + geom_point()+geom_errorbar() + theme(axis.text.x = element_text(angle=90)) + scale_y_sqrt() 
```
Indicates role between seller_zip_code and number of hours for delivery

@@

```{r Explore customer_zip_code_prefix with number_of_hours, echo=FALSE}
# TODO anyway to add regularization or remove the low impact customer_zip_code_prefix from the model
# let's plot errorbar for number of hours vs customer_zip_code_prefix
customers_zip_of_interest <- wrangled_data %>% dplyr::count(customer_zip_code_prefix) %>% filter(n>=5) %>% pull(customer_zip_code_prefix)

wrangled_data %>% filter(customer_zip_code_prefix %in% customers_zip_of_interest) %>% group_by(customer_zip_code_prefix) %>% summarise(mean=mean(number_of_hours),se=sd(number_of_hours)/sqrt(n())) %>% mutate(customer_zip_code_prefix=reorder(customer_zip_code_prefix,mean)) %>%  ggplot(aes(x=customer_zip_code_prefix,y=mean,ymax=mean+2*se,ymin=mean-2*se)) + geom_point()+geom_errorbar() + theme(axis.text.x = element_text(angle=90)) + scale_y_sqrt()
```
customer_zip_code_prefix can  be seen to be  influence number of hours for delivery

@@

```{r Explore product_id with number_of_hours}
# TODO anyway to add regularization or remove the low impact products from the model
# let's plot errorbar for number of hours vs product_id
products_of_interest <- wrangled_data %>% dplyr::count(product_id)  %>% filter(n>=10) %>% pull(product_id)

wrangled_data %>% filter(product_id %in% products_of_interest) %>% group_by(product_id) %>% summarise(mean=mean(number_of_hours),se=sd(number_of_hours)/sqrt(n())) %>% mutate(product_id=reorder(product_id,mean)) %>%  ggplot(aes(x=product_id,y=mean,ymax=mean+2*se,ymin=mean-2*se)) + geom_point()+geom_errorbar() + theme(axis.text.x = element_blank()) 
```
It influences the number of hours

```{r Explore volume of product group with number_of_hours, echo=FALSE}
# let's plot errorbar for number of hours vs volume of product group
wrangled_data %>% mutate(volume_group=round((product_length_cm*product_width_cm*product_height_cm)/1000)+1) %>% mutate(volume_group=as.factor(volume_group)) %>% group_by(volume_group) %>% summarise(mean=mean(number_of_hours),se=sd(number_of_hours)/sqrt(n())) %>% ggplot(aes(x=volume_group,y=mean,ymax=mean+2*se,ymin=mean-2*se)) + geom_point()+geom_errorbar() + theme(axis.text.x = element_text(angle=90))
```
It does not influencee much the number of hours

However a smooth fit shows some pattern 
```{r smooth fit for volume, echo=FALSE}
wrangled_data %>% mutate(volume=product_length_cm*product_width_cm*product_height_cm) %>% ggplot(aes(volume,number_of_hours)) + geom_smooth()
```

```{r Explore weight of product with number_of_hours}
# let's plot errorbar for number of hours vs weight of product
wrangled_data %>% mutate(weight_group=round((product_weight_g)/1000)+1) %>% mutate(weight_group=as.factor(weight_group)) %>% mutate(weight_group=as.factor(weight_group)) %>% group_by(weight_group) %>% summarise(mean=mean(number_of_hours),se=sd(number_of_hours)/sqrt(n())) %>% ggplot(aes(x=weight_group,y=mean,ymax=mean+2*se,ymin=mean-2*se)) + geom_point()+geom_errorbar() + theme(axis.text.x = element_text(angle=90))
```
Some clues but not strongly correlated.

```{r Smooth fit for product, echo= FALSE}
wrangled_data  %>% ggplot(aes(product_weight_g,number_of_hours)) + geom_smooth()
```

>>>>

```{r Install or load packages,include=FALSE}
# TODO use conditional loading or installation of library
# Remove plotly related code for pdfs
if (!require('dplyr')) install.packages('dplyr'); library('dplyr')
if (!require('lubridate')) install.packages('lubridate'); library('lubridate')
if (!require('tidyverse')) install.packages('tidyverse'); library('tidyverse')
if (!require('geosphere')) install.packages('geosphere'); library('geosphere')
if (!require('caret')) install.packages('caret'); library('caret')
if (!require('randomForest')) install.packages('randomForest'); library('randomForest')
if (!require('plotly')) install.packages('plotly'); library('plotly')
if (!require('gam')) install.packages('gam'); library('gam')
```

@@

# let's combine with products data as well
orders_dataset %>% inner_join(order_items_dataset,by='order_id') %>% inner_join(sellers_dataset,by='seller_id') %>% inner_join(customers_dataset,by='customer_id') %>% inner_join(products_dataset,by='product_id') %>% select(-customer_id,-order_status,-seller_city,-seller_state,-customer_city,-customer_state) %>% head()

# so raw data
raw_data <- orders_dataset %>% 
  inner_join(order_items_dataset,by='order_id') %>%
  inner_join(sellers_dataset,by='seller_id') %>% 
  inner_join(customers_dataset,by='customer_id') %>%  
  inner_join(products_dataset,by='product_id') %>%
  mutate(order_purchase_timestamp=as.numeric(order_purchase_timestamp),
         product_id=as.factor(product_id), 
         seller_id=as.factor(seller_id),
         seller_zip_code_prefix=as.factor(seller_zip_code_prefix),
         customer_unique_id=as.factor(customer_unique_id),
         customer_zip_code_prefix=as.factor(customer_zip_code_prefix),
         product_category_name=as.factor(product_category_name)) %>%
  select(-customer_id,-order_status,-seller_city,-seller_state,-customer_city,-customer_state, -payment_approved_at,-order_id, -order_item_id, -product_id,-seller_zip_code_prefix,-customer_unique_id,-customer_zip_code_prefix)
```

Calculating current RMSE range
```{r Current RMSE, echo = FALSE}
# current RMSE can be calculated by finding number of hours for predicted delivery time given by order_estimated_delivery_date vs actual outcome/number of hours give by order_delivered_customer_date
# However, note that order_estimated_delivery_date has no timestamp, so we will remove timestamp component, round date using floor_date, which means any delivery done during the day say at 2017-10-10 21:25:13 is treated to be done on 2017-10-10

# first we convert it to desired date
# then we covert date to numeric i.e. to convert it to seconds
# then we calculate time_taken_from_order_to_delivery, estimated_time_from_order_to_delivery in seconds
data_for_rmse <- raw_data %>%
  select(order_purchase_timestamp,order_delivered_customer_date,order_estimated_delivery_date) %>%
  mutate(order_delivered_customer_date_low=floor_date(order_delivered_customer_date,unit="day"),
         order_delivered_customer_date_high=ceiling_date(order_delivered_customer_date,unit="day")) %>% 
  mutate(order_delivered_customer_date_low=as.numeric(order_delivered_customer_date_low),
         order_delivered_customer_date_high=as.numeric(order_delivered_customer_date_high),
         order_estimated_delivery_date=as.numeric(order_estimated_delivery_date))  %>%
  mutate(time_taken_from_order_to_delivery_rangemax=order_delivered_customer_date_high-order_purchase_timestamp,
         time_taken_from_order_to_delivery_rangemin=order_delivered_customer_date_low-order_purchase_timestamp,
         estimated_time_from_order_to_delivery=order_estimated_delivery_date-order_purchase_timestamp) %>% 
  mutate(estimated_time_from_order_to_delivery=estimated_time_from_order_to_delivery/3600,
         time_taken_from_order_to_delivery_rangemax = time_taken_from_order_to_delivery_rangemax/3600,
         time_taken_from_order_to_delivery_rangemin = time_taken_from_order_to_delivery_rangemin/3600) %>%
  select(estimated_time_from_order_to_delivery,time_taken_from_order_to_delivery_rangemax,time_taken_from_order_to_delivery_rangemin)

RMSE_value_based_on_max <- RMSE(data_for_rmse$time_taken_from_order_to_delivery_rangemax,data_for_rmse$estimated_time_from_order_to_delivery)
# 359.9466

RMSE_value_based_on_min <- RMSE(data_for_rmse$time_taken_from_order_to_delivery_rangemin,data_for_rmse$estimated_time_from_order_to_delivery)
# 377.9515

library(plotly)
ggplotly(p1)

rm(data_for_rmse)
```

```{r Prepare final wrangled data set, echo=FALSE}
# we will predict number of hours to deliver based on various predictors. This seems more appropriate. Though as a result of this, we won't be able to use night, day time data of purchase which may influence delivery time (say to carrier it is informed only next day)
wrangled_data <- raw_data %>%
  mutate(order_delivered_customer_date=as.numeric(order_delivered_customer_date)) %>%
  mutate(number_of_hours=(order_delivered_customer_date-order_purchase_timestamp)/3600) %>%
  select(-order_purchase_timestamp,-order_delivered_customer_date,-order_estimated_delivery_date)
```

```{r Apply one hot encoding to product_category_name, include = FALSE}
set.seed(1,sample.kind = 'Rounding')
# fullRank parameter to make sure there is no perfect collinearity
product_category_dummyvars <- data.frame(predict(dummyVars(~product_category_name, data=wrangled_data, fullRank = TRUE), newdata = wrangled_data))

wrangled_data <- wrangled_data %>% cbind.data.frame(product_category_dummyvars) %>% select(-product_category_name)

wrangled_data %>% str()
rm(product_category_dummyvars)
```

@@

```{r Cleanup and save required dataset}
save(wrangled_data,file='rda/wrangled_data.rda')
rm(geolocation_dataset, order_items_dataset, orders_dataset, p1, product_category_name_translation, products_dataset, sellers_dataset, sellers_to_ignore, customers_dataset, customers_to_ignore, raw_data)
```

Derived data set should be generated later on observations
- geolocation usage to capture distance
- delay and mean delay in handling to carrier
- number of hours for delivery should be generated later
```{r Method to derive distance data, include = FALSE}
# we will use distGeo from geosphere package to compute distance in km between seller and customer. considered to be highly accurate. if it takes more time, we can consider using distVincentyEllipsoid or distHaversine as we are fine with accuracy of a km.
extract_distance_between_seller_customer <- function(df) { 
  df %>% mutate(distance_between_two_points=round(distGeo(
    matrix(c(geolocation_lng_seller,geolocation_lat_seller),ncol=2),
    matrix(c(geolocation_lng_customer,geolocation_lat_customer),ncol=2)
    )/1000,2
    )) %>% select(-geolocation_lat_customer,-geolocation_lng_customer,-geolocation_lat_seller,-geolocation_lng_seller)
}
```

```{r Method to derive delay in handover to carrier, include = FALSE}
# calculated in hours
get_delay_in_handover_to_carrier_per_seller_id <- function(df) {
  df %>% mutate(delay= (as.numeric(handed_over_to_carrier_date) -as.numeric(shipping_limit_date))/3600)  %>% 
    group_by(seller_id) %>% 
    summarise(mean_penalty_to_consider=mean(delay))
}
```

```{r Method to compute volume from available fields, include = FALSE}
# compute volume
compute_volume_using_product_info <- function(df) {
  df %>% mutate(volume=product_length_cm*product_height_cm*product_width_cm) %>%
    select(-product_length_cm,-product_height_cm,-product_width_cm)
}
```

Split data into training and validation set, where validation set will be used later
```{r Split data into training and validation set, include = FALSE}
set.seed(1,sample.kind = 'Rounding')
test_indices <- createDataPartition(wrangled_data$number_of_hours,times=1,p=0.2,list=FALSE)

# ensure validation_dataset has products only for concerned product categories and seller_ids which are in training_dataset 
# validation_dataset <- validation_dataset %>% semi_join(training_dataset,by='product_category_name') %>% semi_join(training_dataset,by='seller_id')
validation_dataset <- validation_dataset %>% semi_join(training_dataset,by='seller_id')

We use training_dataset and split into into train and test set. So that training_dataset is sufficient to fine tune the model
```{r Split data for training, include=FALSE}
# Let's split training_dataset into two parts
set.seed(1,sample.kind = 'Rounding')
training_dataset <- training_dataset %>% compute_volume_using_product_info() %>% extract_distance_between_seller_customer()

train_subset <- training_dataset[-test_indices,]
test_subset <- training_dataset[test_indices,]
# ensure test_subset has products only for concerned product categories and seller_ids which are in train_subset
# test_subset <- test_subset %>% semi_join(train_subset,by='product_category_name') %>% semi_join(train_subset,by='seller_id')
test_subset <- test_subset %>% semi_join(train_subset,by='seller_id')

penalty_data <- train_subset %>% get_delay_in_handover_to_carrier_per_seller_id()

@@

```{r Explore product_category_name relation with number_of_hours, echo=FALSE}
# TODO review it, fix it as required
training_dataset %>% group_by(product_category_name) %>% summarise(mean=mean(number_of_hours),se=sd(number_of_hours)/sqrt(n())) %>% mutate(product_category_name=reorder(product_category_name,mean)) %>% ggplot(aes(x=product_category_name,y=mean,ymax=mean+2*se,ymin=mean-2*se)) + geom_point()+geom_errorbar() + theme(axis.text.x = element_text(angle=90))
```

@@

```{r Explore price relation with number_of_hours, echo = FALSE}
# first let see how are the items associated with orders priced. Most of the order items a
training_dataset %>% mutate(price_group=round(price/100)+1) %>% ggplot(aes(price_group)) + geom_histogram(binwidth=1) + xlab('Each price_group is of 100 units each')

# let's plot errorbar for number of hours vs price_group.
# we can see while for price groups number of hours for delivery vary but the relationship is not clear enough. Higher priced items can also be delivered early
training_dataset %>% mutate(price_group=round(price/50)+1) %>% mutate(price_group=as.factor(price_group)) %>% group_by(price_group) %>% summarise(mean=mean(number_of_hours),se=sd(number_of_hours)/sqrt(n())) %>% ggplot(aes(x=price_group,y=mean,ymax=mean+2*se,ymin=mean-2*se)) + geom_point()+geom_errorbar() + theme(axis.text.x = element_text(angle=90))

cor(training_dataset$price,training_dataset$number_of_hours)
```

@@

```{r Explore distance relationship between two points with number_of_hours, echo = FALSE}
# let's plot errorbar for number of hours vs distance group
# we can see a pattern that as distance increases delivery time increases. 
training_dataset %>% mutate(distance_group=round(distance_between_two_points/50)+1) %>% mutate(distance_group=as.factor(distance_group)) %>% group_by(distance_group) %>% summarise(mean=mean(number_of_hours),se=sd(number_of_hours)/sqrt(n())) %>% ggplot(aes(x=distance_group,y=mean,ymax=mean+2*se,ymin=mean-2*se)) + geom_point()+geom_errorbar() + theme(axis.text.x = element_text(angle=90))

cor(training_dataset$distance_between_two_points,training_dataset$number_of_hours)
```

@@

```{r Explore freight value group with number_of_hours}
# let's plot errorbar for number of hours vs freight value group
training_dataset %>% mutate(freight_value_group=round(freight_value/5)+1) %>% mutate(freight_value_group=as.factor(freight_value_group)) %>% group_by(freight_value_group) %>% summarise(mean=mean(number_of_hours),se=sd(number_of_hours)/sqrt(n())) %>% ggplot(aes(x=freight_value_group,y=mean,ymax=mean+2*se,ymin=mean-2*se)) + geom_point()+geom_errorbar() + theme(axis.text.x = element_text(angle=90))

cor(training_dataset$freight_value,training_dataset$number_of_hours)
```

@@

However, it dips near the end
```{r smooth fit for freight value, echo=FALSE}
training_dataset %>% ggplot(aes(x=freight_value,y=number_of_hours)) + geom_smooth() 
```

```{r Explore volume of product group with number_of_hours, echo=FALSE}
# let's plot errorbar for number of hours vs volume of product group
training_dataset %>% mutate(volume_group=round(volume/1000)+1) %>% mutate(volume_group=as.factor(volume_group)) %>% group_by(volume_group) %>% summarise(mean=mean(number_of_hours),se=sd(number_of_hours)/sqrt(n())) %>% ggplot(aes(x=volume_group,y=mean,ymax=mean+2*se,ymin=mean-2*se)) + geom_point()+geom_errorbar() + theme(axis.text.x = element_text(angle=90))
```
It does not influencee much the number of hours

However a smooth fit shows some pattern 
```{r smooth fit for volume, echo=FALSE}
training_dataset %>% ggplot(aes(volume,number_of_hours)) + geom_smooth()
```

```{r Explore weight of product with number_of_hours}
# let's plot errorbar for number of hours vs weight of product
training_dataset %>% mutate(weight_group=round((product_weight_g)/1000)+1) %>% mutate(weight_group=as.factor(weight_group)) %>% mutate(weight_group=as.factor(weight_group)) %>% group_by(weight_group) %>% summarise(mean=mean(number_of_hours),se=sd(number_of_hours)/sqrt(n())) %>% ggplot(aes(x=weight_group,y=mean,ymax=mean+2*se,ymin=mean-2*se)) + geom_point()+geom_errorbar() + theme(axis.text.x = element_text(angle=90))
```
Some clues but not strongly correlated.

```{r Smooth fit for product, echo= FALSE}
training_dataset  %>% ggplot(aes(product_weight_g,number_of_hours)) + geom_smooth()
```

```{r Prepare training and test data for ,include=FALSE}
# penalty_data is computed on train_subset and is used with test_subset as historical learning of efficiency of each seller in delivering to carrier
train_subset <- train_subset %>% left_join(penalty_data,by='seller_id') %>% select(-handed_over_to_carrier_date, -shipping_limit_date, -seller_id)

test_subset <- test_subset %>% left_join(penalty_data,by='seller_id') %>% select(-handed_over_to_carrier_date, -shipping_limit_date, -seller_id)
```

```{r knn model, echo = FALSE}
set.seed(seed=1,sample.kind = 'Rounding')

tuningControl=trainControl(method = "cv", number = 3, p = .8)
knn_fit <- train(number_of_hours~., train_subset, method='knn',tuneGrid=data.frame(k=c(27,34,41)),trControl=tuningControl)

RMSE(predictions,(test_subset %>% pull(number_of_hours) ))
#199.448 for k=34

```{r knn model data scaline under test, echo=FALSE}
# TODO review and improve or remove this as required
set.seed(seed=1,sample.kind = 'Rounding')
scaled_train_subset <- train_subset %>% mutate(across(c(mean_penalty_to_consider,distance_between_two_points,volume,product_weight_g,freight_value,price),scale)) 
scaled_test_subset <- test_subset %>% mutate(across(c(mean_penalty_to_consider,distance_between_two_points,volume,product_weight_g,freight_value,price),scale))

tuningControl=trainControl(method = "cv", number = 3, p = .8)
knn_fit <- train(number_of_hours~., scaled_train_subset, method='knn',tuneGrid=data.frame(k=c(27,34,41)),trControl=tuningControl)

predictions <- predict(knn_fit,scaled_test_subset %>% select(-number_of_hours))

RMSE(predictions,(scaled_test_subset %>% pull(number_of_hours) ))
#194.8233 for k=34

```{r rf model, echo = FALSE}
set.seed(seed=1,sample.kind = 'Rounding')
tuningControl=trainControl(method = "cv", number = 2, p = .8)
rf_fit <- train(number_of_hours~., train_subset, method='rf',tuneGrid = data.frame(mtry = seq(5, 50, 10)),trControl=tuningControl,ntree=250) # nodesize = 50,maxnodes=2000,ntree=200

RMSE(predictions,(test_subset %>% pull(number_of_hours) ))
# 184.1709 for mtry 15. Technically there are total of 79 variables due to 73 product categories and 6 other predictors

RMSE(predictions,(test_subset %>% pull(number_of_hours) ))
# 195.3235

RMSE(predictions,(test_subset %>% pull(number_of_hours) ))
# 198.8575 (improves if we use seller_id as predictor as well)
```
Warning:
prediction from a rank-deficient fit may be misleading

RMSE(predictions,(test_subset %>% pull(number_of_hours) ))
#194.6215

```{r GamLoess model with scaling, echo = FALSE}
# TODO review-keep or remove it
set.seed(seed=1,sample.kind = 'Rounding')
scaled_train_subset <- train_subset %>% mutate(across(c(mean_penalty_to_consider,distance_between_two_points,volume,product_weight_g,freight_value,price),scale)) 
scaled_test_subset <- test_subset %>% mutate(across(c(mean_penalty_to_consider,distance_between_two_points,volume,product_weight_g,freight_value,price),scale))

tuningControl=trainControl(method = "cv", number = 3, p = .8)
#grid <- expand.grid(span = seq(0.15, 0.65, len = 10), degree = 1)
grid <- expand.grid(span = seq(0.15, 0.65, len = 3), degree = 1)
gamLoess_fit <- train(number_of_hours~., scaled_train_subset, method='gamLoess',tuneGrid=grid,trControl=tuningControl)

predictions <- predict(gamLoess_fit,scaled_test_subset %>% select(-number_of_hours))

RMSE(predictions,(scaled_test_subset %>% pull(number_of_hours) ))
#194.7017

save(gamLoess_fit,file='rda/gamLoess_model.rda')
```
Warnings received:
extrapolation not allowed with blending
 (Discovered by lowesd)liv too small.    
 (Discovered by lowesd)lv too small
 extrapolation not allowed with blending
 non-list contrasts argument ignored

```{r Final Model Training, echo = FALSE}
set.seed(seed=1,sample.kind = 'Rounding')

penalty_data <- training_dataset %>% get_delay_in_handover_to_carrier_per_seller_id()

training_dataset <- training_dataset %>% left_join(penalty_data,by='seller_id') %>% select(-seller_id,-handed_over_to_carrier_date,-shipping_limit_date)

mtry_to_use <- rf_fit$finalModel$mtry
ntree_to_use <- 500 #rf_fit$finalModel$ntree
tuningControl <- trainControl(method = "cv", number = 2, p = .8)
final_model <- train(number_of_hours~., training_dataset, method='rf',tuneGrid = data.frame(mtry = c(mtry_to_use)),trControl=tuningControl,ntree=ntree_to_use)

# final_model <- randomForest(number_of_hours~., training_dataset,mtry= mtry_to_use, ntree=ntree_to_use)

final_model

save(final_model,file='rda/final_model.rda')
```

```{r Final movel validation, echo = FALSE}
validation_dataset <- validation_dataset %>% compute_volume_using_product_info() %>% extract_distance_between_seller_customer() %>% left_join(penalty_data,by='seller_id') %>% select(-seller_id,-handed_over_to_carrier_date,-shipping_limit_date)

predictions <- predict(final_model,validation_dataset %>% select(-number_of_hours))

RMSE((validation_dataset %>% pull(number_of_hours) ),predictions)
# 194.6748 for mtry=25, ntree=500
```

===== TODO review from here

Data used maybe incorrect
Review joins and if multiple orders of same type are being delivered at the same time
refer
> training_dataset %>% filter(number_of_hours>=354.73 & number_of_hours<354.74)
   price freight_value product_category_name product_weight_g number_of_hours volume
1   59.0         13.43          garden_tools             1550        354.7386  19800
2   59.0         13.43          garden_tools             1550        354.7386  19800
3   59.0         13.43          garden_tools             1550        354.7386  19800
4   59.0         13.43          garden_tools             1550        354.7386  19800
5   59.0         13.43          garden_tools             1550        354.7386  19800
6   59.0         13.43          garden_tools             1550        354.7386  19800
7   59.0         13.43          garden_tools             1550        354.7386  19800
8   59.0         13.43          garden_tools             1550        354.7386  19800
9   59.0         13.43          garden_tools             1550        354.7386  19800
10  59.0         13.43          garden_tools             1550        354.7386  19800
11  59.0         13.43          garden_tools             1550        354.7386  19800
12  66.0         14.21       furniture_decor             1150        354.7319  28175
13  46.9         17.63        sports_leisure              300        354.7378   2304
   distance_between_two_points mean_penalty_to_consider
1                       469.74                -51.60514
2                       469.74                -51.60514
3                       469.74                -51.60514
4                       469.74                -51.60514
5                       469.74                -51.60514
6                       469.74                -51.60514
7                       469.74                -51.60514
8                       469.74                -51.60514
9                       469.74                -51.60514
10                      469.74                -51.60514
11                      469.74                -51.60514
12                      807.32                -56.57523
13                     1083.46               -147.48749

```{r Derived data}
# seller to customer distance should be right predictor to consider, as the text says "Those merchants are able to sell their products through the Olist Store and ship them directly to the customers using Olist logistics partners."
# source: https://www.kaggle.com/olistbr/brazilian-ecommerce
wrangled_data %>% head() %>% extract_distance_between_seller_customer()

# plot of how 
p1 <- wrangled_data %>% extract_distance_between_seller_customer() %>% ggplot(aes(distance_between_two_points)) + geom_histogram(binwidth=50)
ggplotly(p1)
```

```{r Explore delayed handover to carrier, echo=FALSE }
# TODO we maybe able to remove this as it is not used, because this is the info which is generally not available
# identify if there was delay in delivery to carrier
wrangled_data <- wrangled_data %>% mutate(delayed_to_carrier=if_else((as.numeric(handed_over_to_carrier_date)-as.numeric(shipping_limit_date))>0,TRUE,FALSE))

wrangled_data <- wrangled_data %>% select(-shipping_limit_date,-handed_over_to_carrier_date,-order_estimated_delivery_date)
```

@@

```{r Explore seller_zip_code_prefix with number_of_hours, echo = FALSE}
# TODO anyway to add regularization or remove the low impact seller_zip_code_prefix from the model
# let's plot errorbar for number of hours vs seller_zip_code_prefix
sellers_zip_of_interest <- training_dataset %>% dplyr::count(seller_zip_code_prefix) %>% filter(n>10) %>% pull(seller_zip_code_prefix)

training_dataset %>% filter(seller_zip_code_prefix %in% sellers_zip_of_interest) %>% group_by(seller_zip_code_prefix) %>% summarise(mean=mean(number_of_hours),se=sd(number_of_hours)/sqrt(n())) %>% mutate(seller_zip_code_prefix=reorder(seller_zip_code_prefix,mean)) %>%  ggplot(aes(x=seller_zip_code_prefix,y=mean,ymax=mean+2*se,ymin=mean-2*se)) + geom_point()+geom_errorbar() + theme(axis.text.x = element_text(angle=90)) + scale_y_sqrt() 
```
Indicates role between seller_zip_code and number of hours for delivery

@@

```{r Explore customer_zip_code_prefix with number_of_hours, echo=FALSE}
# TODO anyway to add regularization or remove the low impact customer_zip_code_prefix from the model
# let's plot errorbar for number of hours vs customer_zip_code_prefix
customers_zip_of_interest <- training_dataset %>% dplyr::count(customer_zip_code_prefix) %>% filter(n>=5) %>% pull(customer_zip_code_prefix)

training_dataset %>% filter(customer_zip_code_prefix %in% customers_zip_of_interest) %>% group_by(customer_zip_code_prefix) %>% summarise(mean=mean(number_of_hours),se=sd(number_of_hours)/sqrt(n())) %>% mutate(customer_zip_code_prefix=reorder(customer_zip_code_prefix,mean)) %>%  ggplot(aes(x=customer_zip_code_prefix,y=mean,ymax=mean+2*se,ymin=mean-2*se)) + geom_point()+geom_errorbar() + theme(axis.text.x = element_text(angle=90)) + scale_y_sqrt()
```
customer_zip_code_prefix can  be seen to be  influence number of hours for delivery

@@

```{r Explore product_id with number_of_hours}
# TODO anyway to add regularization or remove the low impact products from the model
# let's plot errorbar for number of hours vs product_id
products_of_interest <- training_dataset %>% dplyr::count(product_id)  %>% filter(n>=10) %>% pull(product_id)

training_dataset %>% filter(product_id %in% products_of_interest) %>% group_by(product_id) %>% summarise(mean=mean(number_of_hours),se=sd(number_of_hours)/sqrt(n())) %>% mutate(product_id=reorder(product_id,mean)) %>%  ggplot(aes(x=product_id,y=mean,ymax=mean+2*se,ymin=mean-2*se)) + geom_point()+geom_errorbar() + theme(axis.text.x = element_blank()) 
```
It influences the number of hours

====TODO review above
